State of the Practice for Medical Imaging Software

### 2022-04-13 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Edits to SOP for MI paper

The frequency of checkmarks in Table~\ref{Tbl_Guidelines} indicates the
popularity of recommending a given artifact, but it does not imply that the most
commonly recommended artifacts are the most important artifacts. Just because an
artifact is not explicitly recommended in a given guidelines, does not mean that
the artifact is not valued by the guideline authors.  They may have exluded it
because it is out of the scope of their recommendations, or outside of their
experience.  For instance, an artifact related to uninstall is only explicitly
mentioned by \citep{vanGompelEtAl2016}), but other guideline authors would
likely see its value.  They may simply feel that uninstall is implied by
install, or they may have never asked themselves whether separate uninstall
instructions are needed.

@@

Although the MI community shows examples of 88\% (23 of 26) of the practices we
found in research software guidelines (Table~\ref{Tbl_Guidelines}), three
recommended artifacts were not observed: i) Unistall, ii) Test plans, and iii)
Requirements.  Uninstall is likely an omission caused by the focus on installing
software. Given the storage capacity of current hardware, developers and users
are not generally concerned with uninstall.  Moreover, as mentioned above,
uninstall is not particularly emphasized in existing recommendations.  Test
plans were not observed for MI software, but that doesn't mean they weren't
created; it means that the plans are not under version control.  Test plans
would have to at least be implicitly created, since test cases were observed
with reasonable frequency for MI software (test cases are are categorized as
uncommon).

@@

Table~\ref{Tbl_Guidelines} shows several artifacts that are rarely observed in
practice.  A theme among these rare artifacts is that many of them are related
to developers more than users.  For instance, the rare artifacts include
Contributing, Developer code of conduct, Code style guide, Product roadmap,
Requirements, Design documentation, API documentation and a Test plan.  The rare
artifacts for MI software are similar to the rare artifacts for Lattice
Boltzmann solvers \citep{Michalski2021}, except LBM software is more likely to
have developer related artifacts, like Contributing, Dependency list, and Design
documentation.

To improve MI software in the future, an increased use of checklists could help.
Checklists can be used in projects to ensure that best practices are followed by
all developers.  Some examples include checklists merging branches into master
\citep{Brown2015}, checklists for saving and sharing changes to the project
\citep{WilsonEtAl2016}, checklists for new and departing team members
\citep{HerouxAndBernholdt2018}, checklists for processes related to commits and
releases \citep{HerouxEtAl2008} and checklists for overall software quality
\citep{ThielEtAl2020, SSI2022}.  For instace, for Lattice Boltzmann solver
software, ESPResSo has a checklist for managing releases \citep{Michalski2021}.

@@

>>>>

The frequency of checkmarks in Table~\ref{Tbl_Guidelines} indicates the
popularity of recommending a given artifact, but it does not imply that the most
commonly recommended artifacts are the most important artifacts. Just because an
artifact is not explicitly recommended in a given guidelines, does not mean that
the artifact is not valued by the guideline authors.  They may have excluded it
because it is out of the scope of their recommendations, or outside of their
experience.  For instance, an artifact related to uninstall is only explicitly
mentioned by \citep{vanGompelEtAl2016}), but other guideline authors would
likely see its value.  They may simply feel that uninstall is implied by
install, or they may have never asked themselves whether separate uninstall
instructions are needed.

@@

Although the MI community shows examples of 88\% (23 of 26) of the practices we
found in research software guidelines (Table~\ref{Tbl_Guidelines}), three
recommended artifacts were not observed: i) Uninstall, ii) Test plans, and iii)
Requirements.  Uninstall is likely an omission caused by the focus on installing
software. Given the storage capacity of current hardware, developers and users
are not generally concerned with uninstall.  Moreover, as mentioned above,
uninstall is not particularly emphasized in existing recommendations.  Test
plans were not observed for MI software, but that doesn't mean they weren't
created; it means that the plans are not under version control.  Test plans
would have to at least be implicitly created, since test cases were observed
with reasonable frequency for MI software (test cases are are categorized as
uncommon).

@@

Table~\ref{Tbl_Guidelines} shows several artifacts that are rarely observed in
practice.  A theme among these rare artifacts is that many of them are related
to developers more than users.  For instance, the rare artifacts include
Contributing, Developer code of conduct, Code style guide, Product roadmap,
Requirements, Design documentation, API documentation and a Test plan.
\wss{additional details are available in the LBM document for other missing
parts, like code style guides.}  The rare artifacts for MI software are similar
to the rare artifacts for Lattice Boltzmann solvers \citep{Michalski2021},
except LBM software is more likely to have developer related artifacts, like
Contributing, Dependency list, and Design documentation.

To improve MI software in the future, an increased use of checklists could help.
Checklists can be used in projects to ensure that best practices are followed by
all developers.  Some examples include checklists merging branches into master
\citep{Brown2015}, checklists for saving and sharing changes to the project
\citep{WilsonEtAl2016}, checklists for new and departing team members
\citep{HerouxAndBernholdt2018}, checklists for processes related to commits and
releases \citep{HerouxEtAl2008} and checklists for overall software quality
\citep{ThielEtAl2020, SSI2022}.  For instance, for Lattice Boltzmann solver
software, ESPResSo has a checklist for managing releases \citep{Michalski2021}.

@@

\citet{PintoEtAl2018} lists some pain points that did not come up in our
conversations with LBM developers: Cross-platform compatibility, interruptions
while coding, scope bloat, lack of user feedback, hard to collaborate on
software projects, and aloneness. \citet{WieseEtAl2019} repeat some of the
previous pain points and add the following: dependency management, data handling
concerns (like data quality, data management and data privacy), reproducibility,
and software scope determination. Although LBM developers did not mention these
pain points, we cannot conclude that they are not relevant for LBM software
development, since we only interviewed four LBM developers for about an hour each.

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
	Development Time} A developer of pyLBM noted that their small development
	team has a lack of time to implement new features. Small development teams
	are common for LBM software packages (as shown in the measurement table
	excerpt in Figure~\ref{measurement_template_image}). Lack of time is also
	highlighted as a pain point by other research software developers
	\citet{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.
